\documentclass[12pt]{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{mathrsfs}


\usepackage{soul}


% Dr. David Brown Commands
\newcommand{\ds}{\displaystyle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}


% Commands to get script r
\usepackage{calligra}
\DeclareMathAlphabet{\mathcalligra}{T1}{calligra}{m}{n}
\DeclareFontShape{T1}{calligra}{m}{n}{<->s*[2.2]callig15}{}
\newcommand{\scripty}[1]{\ensuremath{\mathcalligra{#1}}}

% Unit commands
\newcommand{\degree}{\text{°}}
\newcommand{\unitSpeed}{\frac{\text{m}}{\text{s}}}
\newcommand{\unitAcceleration}{\frac{\text{m}}{\text{s}^2}}

% Constant commands
\newcommand{\avogadro}{6.022\times10^{23}}

\newcommand{\boltzmannConstK}{1.381\times10^{-23}\frac{\text{J}}{\text{K}}}
\newcommand{\boltzmannConstInEV}{8.617\times10^{-5}\frac{\text{eV}}{\text{K}}}
\newcommand{\planckConst}{6.626\times10^{-34}\text{J}\cdot\text{s}}

\newcommand{\atmP}{1.01325\times10^5\,\text{Pa}}

% Commands to easily write partial derivatives
\newcommand{\partDeriv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\doublePartDeriv}[2]{\frac{\partial^2 #1}{\partial^2 #2}}


\title{Problem 6.36}
\author{Gabriel Decker}


\begin{document}



\maketitle
\begin{flushleft}


In this problem, we must calculate the average speed of ideal gases under the Maxwell Speed Distribution.
This means that it follows the assumptions of Boltzmann statistics (the system is in thermal equilibrium with a source at temperature $T$ and all particles are confined to the system).
This yields the following distribution of speeds.
\begin{equation}
    \mathcal{D}(v)=4\pi\left(\frac{m}{2\pi kT}\right)^{3/2}v^2e^{-mv^2/2kT}
\end{equation}
To calculate the probability of a range of speeds, we use the following.
\begin{equation}
    \mathcal{P}([v_1,v_2])=\int_{v_1}^{v_2}\mathcal{D}(v)\,dv
\end{equation}
    To calculate the probability of a particular speed (with an incredibly small displacement), we use the following.
\begin{equation}
    \mathcal{P}(v)=\mathcal{D}(v)\,dv
\end{equation}\\~\\

To calculate the average speed, we will begin with the following equation.
\begin{equation}
    \overline{v}=\sum_\text{all v}v\mathcal{D}(v)\,dv
    \label{eq:avgSpeed}
\end{equation}
This comes from the following, where the average of a quantity is simply the sum of all the possible quantities with the probability of that quantity existing.
\begin{equation}
    \overline{X}=\sum_sX(s)\mathcal{P}(s)
\end{equation}\\~\\

Beginning with equation \ref{eq:avgSpeed}, we will first assume that there are an infinite many infinitesimal ranges of speeds. This allows us to replace the summation with an integral from $v=0$ to $v=\infty$ (since speed can't be negative).
$$\overline{v}=\sum_\text{all v}v\mathcal{D}(v)\,dv$$
$$\implies\overline{v}=\int_0^\infty v\mathcal{D}(v)\,dv$$
Plugging in $\mathcal{D}(v)$, we get
$$\implies\overline{v}=\int_0^\infty 4\pi\left(\frac{m}{2\pi kT}\right)^{3/2}v^3e^{-mv^2/2kT}\,dv$$
$$\implies\overline{v}=4\pi\left(\frac{m}{2\pi kT}\right)^{3/2}\int_0^\infty v^3e^{-mv^2/2kT}\,dv$$
Substituting $a=m/2kT$, we get
$$\implies\overline{v}=4\pi\left(\frac{a}{\pi}\right)^{3/2}\int_0^\infty v^3e^{-av^2}\,dv$$\\~\\

This is a standard integral with the solution shown by the included script: quick\_integration.py to result in the following.
$$\implies\overline{v}=4\pi\left(\frac{a}{\pi}\right)^{3/2}\cdot\frac{1}{2a^2}$$
$$\implies\overline{v}=2\frac{a^{3/2}}{\pi^{3/2}}\cdot\frac{\pi}{a^2}$$
$$\implies\overline{v}=2a^{3/2}\cdot a^{-4/2}\cdot\pi^{2/2}\cdot\pi^{-3/2}$$
$$\implies\overline{v}=2a^{-1/2}\cdot\pi^{-1/2}$$
$$\implies\overline{v}=2\sqrt{\frac{2kT}{\pi m}}$$
$$\implies\overline{v}=\sqrt{\frac{8kT}{\pi m}}$$



\end{flushleft}

\end{document}
